Designing Types for R, Empirically

### 2020-05-15 paper/main.tex
Add table for uni-typed values

\subsection{Is our simple type language expressive enough to type R code?} Functions in
R can support arguments of multiple types by reflecting on the type to
perform appropriate conversions and computations. On the one hand, this form
of ad-hoc polymorphism makes R functions expressive; on the other hand, it
requires support for polymorphism in the typing scheme, making the type
system complex.  Our system lacks explicit support for
polymorphism. Instead, it supports union, any and ... types. A union type is
inferred when values of multiple types are observed at a function parameter
and return position, any type is inferred when an argument is missing or
unused, and ... is inferred for vararg parameters. Furthermore, list element
types can be inferred as any if they have more than 5 different types. Values with many different types can result in
a large number of alternatives in the type inferred for that position,
making it hard to read and unwieldy for end-users. To evaluate this
''expressivity gap`' between our type inference scheme and R functions, we
look at the number of different types in the parameter and return types of
the XXX functions in our corpus.

@@

We observe that XXX% of the positions have a single type ascribed to them. XXX%
have five or more types ascribed to them. Parameter PARAMNAME of function
XXXXCOLONCOLONXXX has the maximum, XXX, types ascribed to it. This implies that
most functions admit or return a uni-typed value.

>>>>

\subsection{Is our simple type language expressive enough to type R code?} Functions in
R can support arguments of multiple types by reflecting on the type to
perform appropriate conversions and computations. On the one hand, this form
of ad-hoc polymorphism makes R functions expressive; on the other hand, it
requires support for polymorphism in the typing scheme, making the type
system complex.  Our system lacks explicit support for
polymorphism. Instead, it supports union, \any and \VARGS types. A union type is
inferred when values of multiple types are observed at a function parameter
and return position, \any type is inferred when an argument is missing or
unused, and \VARGS is inferred for variable length argument list positions. Furthermore, list element
types can be inferred as \any if they have more than 5 different types. Values with many different types can result in
a large number of alternatives in the type inferred for that position,
making it hard to read and unwieldy for end-users. To evaluate this
''expressivity gap`' between our type inference scheme and R functions, we
look at the number of different types in the parameter and return types of
the XXX functions in our corpus.

@@

We observe that XXX of the positions have a single type ascribed to them, and,
only XXX have four or more types ascribed to them. Parameter PARAMNAME of function
XXXXCOLONCOLONXXX has the maximum, XXX, types ascribed to it.
Hence, most functions admit or return a uni-typed value.

Figure 1 provides a breakdown of the different type categories observed at
uni-typed positions.

\input{tables/uni-typed-breakdown.tex}
